Untargeted control doc

### 2023-09-26 untargeted/paruntargeted_control.Rmd
Switching to fully-manual Excel annotation

stan_mix_data <- peak_data %>% 
  select(feature, filename, into) %>%
  left_join(parametadata) %>%
  filter(samp_type=="Std") %>%
  filter(!str_detect(filename, "Mat")) %>%
  mutate(mix=str_extract(filename, "Mix\\d")) %>%
  select(feature, mix, into) %>%
  pivot_wider(names_from = mix, values_from = into, values_fn = mean)

stan_annotations <- best_feat_data %>%
  pmap(function(...){
    row_data <- list(...)
    all_stans %>%
      select(compound_name, mz, rt) %>%
      filter(mz%between%pmppm(row_data$mzmed, 10)) %>%
      mutate(feature=row_data$feature, mzmed=row_data$mzmed,
             rtmed=row_data$rtmed, n=row_data$n)
  }, .progress = TRUE) %>%
  bind_rows() %>%
  arrange(mz, rt, rtmed) %>%
  left_join(stan_mix_data) %>%
  left_join(all_stans %>% select(compound_name, mix)) %>%
  mutate(mix_ratio=ifelse(mix=="Mix1", Mix1/Mix2, Mix2/Mix1)) %>%
  filter(mix_ratio>3 | is.na(mix_ratio)) %>%
  # select(feature, compound_name)
  ungroup()

missed_stans <- all_stans %>%
  anti_join(stan_annotations %>% select(compound_name)) %>%
  print(n=Inf)

stan_choose <- function(df, compound_name_i, feature_i){
  anti_df <- df %>%
    filter(compound_name==compound_name_i) %>%
    filter(feature!=feature_i)
  join_vec <- join_by(compound_name, mz, rt, feature, mzmed, rtmed, n, 
                      Mix1, Mix2, mix, mix_ratio)
  df %>% anti_join(anti_df, by = join_vec)
}
stan_add <- function(df, compound_name_i, feature_i){
  add_row(df, compound_name=compound_name_i, feature=feature_i)
}
stan_remove <- function(df, compound_name_i){
  df %>% filter(!compound_name%in%compound_name_i)
}
if(amendment=="Amm"){
  if(polarity=="pos"){
    stan_annotations %>%
      stan_choose("Sarcosine", "FT0348") %>%
      stan_choose("beta-Alanine", "FT0347") %>%
      stan_choose("Cytosine", "FT0769") %>%
      stan_add("Cytosine from cytidine", "FT0770") %>%
      stan_choose("L-Proline", "FT0862") %>%
      stan_remove("5-Oxoproline") %>%
      stan_choose("L-Leucine", "FT1244") %>%
      stan_choose("L-Isoleucine", "FT1248") %>%
      stan_choose("Proline betaine", "FT1572") %>%
      stan_choose("Betonicine", "FT1842") %>%
      stan_choose("Carnitine", "FT1857") %>%
      stan_remove("Dimethylglycine") %>%
      stan_remove("4-Aminobutyric acid") %>%
      stan_remove("Creatinine from creatine?") %>%
      stan_remove("L-Valine") %>%
      stan_remove(c("L-Threonine", "L-Homoserine")) %>%
      stan_add("Threonine/Homoserine", "FT0953") %>%
      # stan_remove("Allopurinol") %>%
      stan_remove("Guanine from guanosine?") %>%
      stan_add("L-Glutamic acid", "FT1635") %>%
      group_by(feature) %>%
      group_by(compound_name) %>%
      # filter(n()>1)
      ungroup() -> stan_annotations
  } else if(polarity=="neg"){

  } else {
    stop("Unknown polarity")
  }
}
if(amendment=="Nit"){
  if(polarity=="pos"){

mass_check(c("Cytidine"=244.0933))
```

>>>>

pdf(paste0(output_folder, "anno_eics.pdf"), width = 8, height = 4)
all_stans %>%
  filter(compound_type!="Internal Standard") %>%
  mutate(mz=ifelse(compound_name=="Carnitine", mz[which(compound_name=="Threonine betaine?")], mz)) %>%
  group_by(mz) %>%
  summarise(cmpd_lab=paste0(compound_name, " (", mix, ",", rt, ")", collapse = ", ")) %>%
  # head(6) %>%
  pwalk(function(...){
    row_data <- list(...)
    mass_feats <- best_feat_data %>% filter(mzmed%between%pmppm(row_data$mz, 5))
    chosen_eic <- open_dataset("tmzMLs/pqds") %>%
      filter(polarity==!!polarity) %>%
      filter(amendment==!!amendment) %>%
      filter(samp_type=="Poo" | samp_type=="Std") %>%
      filter(mz%between%pmppm(row_data$mz, 5)) %>%
      filter(rt%between%c(2, 20)) %>%
      dplyr::collect()
    gp_plot <- chosen_eic %>%
      left_join(parametadata, by = join_by(filename, samp_type, amendment, polarity)) %>%
      mutate(samp_type=str_extract(filename, "Mix\\d|Deep|Surf|Poo|Blk")) %>%
      filter(!str_detect(filename, "Mat")) %>%
      mutate(plot_facet=ifelse(samp_type%in%c("Mix1", "Mix2"), "Mixes", "Samples")) %>%
      ggplot() +
      geom_line(aes(x=rt, y=int, group=filename, color=samp_type)) +
      facet_wrap(~plot_facet, scales = "free_y", ncol=1) +
      geom_vline(aes(xintercept=rtmed), color="black", data = mass_feats) +
      geom_text(aes(x=rtmed, y=Inf, label=feature), 
                angle=90, hjust=1, vjust=0.5, data = mass_feats) +
      ggtitle(row_data$cmpd_lab) +
      scale_x_continuous(breaks = 2:20) +
      theme_bw() +
      theme(legend.position = c(0.9, 0.8), legend.title = element_blank(),
            axis.title = element_blank())
    print(gp_plot)
  }, .progress = TRUE)
dev.off()
```
